---
title: "Chopped Episodes"
author: "Ted Laderas"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Look at the available datasets

```{r eval=FALSE}
library(tidytuesdayR)
#This will open up in the help window
tidytuesdayR::tt_available()
```

# What was your dataset?

Load your dataset in with the function below. The input is the date the dataset was issued. You should be able to get this from the `tt_available()` function.

```{r}
#incoming data comes in as a list
datasets <- tidytuesdayR::tt_load("2020-08-25")
#show the names of the individual datasets
names(datasets)
```

```{r}
chopped <- datasets$chopped
```

# Initial EDA

```{r}
visdat::vis_dat(chopped)
```

```{r}
skimr::skim(chopped)
```

# What was your question?

Given your inital exploration of the data, what was the question you wanted to answer?

> Which Judge is associated with the highest ratings?

## Fixing the data

Need to recode some names here. Also collapse `judge1`, `judge2` and `judge3` into a single column using `pivot_longer()`.

```{r}
chopped_collapsed <- chopped %>%
  tidyr::pivot_longer(cols=contains("judge"), names_to="judge", values_to="name") %>%
  mutate(name=fct_recode(name, `Aarón Sanchez`="Aarón Sánchez", `Amanda Freitag` = "Amanda Freita")) 
```

## Frequency of Judge appearances

Who is the most frequent judge on Chopped?

```{r}
chopped_counts <- chopped_collapsed %>% 
  count(name) %>%
  arrange(desc(n))

knitr::kable(chopped_counts)
```

## Boxplot of Judge versus Ratings

```{r}
chopped_counts_high <- chopped_counts %>% filter(n > 4) 

frequent_judges <- chopped_counts_high %>% pull(name)

chopped_collapsed %>%
  tidyr::drop_na(episode_rating) %>%
  filter(name %in% frequent_judges) %>%
  mutate(name=fct_reorder(name, episode_rating, median)) %>%
  ggplot() +
  aes(x=name, y=episode_rating, fill=name) +
  geom_boxplot() + 
  geom_text(aes(x = name, y= 10, label=n), data=chopped_counts_high) +
  ylim(c(5.5, 10.5)) +
  theme(legend.position = "none") +
  coord_flip() +
  labs(title = "Chopped Judges by Median Rating", 
       subtitle = "Alton and Tiffani FTW")
```

## Alex's Progression

Alex Guarnaschelli is the most frequent judge. Here's her overall progression over episodes.

```{r}
chopped_collapsed %>%
  filter(name== "Alex Guarnaschelli") %>%
  ggplot() +
  aes(x=series_episode, y=episode_rating, color=season) +
  geom_point() +
  geom_smooth()
```

## Judges versus ratings

Doing something similar for those judges who have 16 or more appearances:


```{r}
frequent_judges <-  chopped_counts %>% 
  filter(n > 15) %>% 
  pull(name)

chopped_collapsed %>%
  filter(name %in% frequent_judges) %>%
  ggplot() +
  aes(x=series_episode, y=episode_rating) +
  geom_point() +
  geom_smooth(aes(color=name)) +
  facet_wrap(~name) +
  ylim(c(6, 9.5)) +
  theme(legend.position = "none")
```